%% Template: IEEE Bare Demo Template for Conferences
%% https://www.overleaf.com/latex/templates/ieee-bare-demo-template-for-conferences/ypypvwjmvtdf
\documentclass[conference]{IEEEtran}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{bm}

\begin{document}
\title{\huge Metaheuristic Solver PRINTEMPS \\for MaxSAT Evaluation 2026}

% author names and affiliations
\author{\IEEEauthorblockN{Yuji Koguma}
\IEEEauthorblockA{
Email: yuji.koguma@gmail.com}
\and
\IEEEauthorblockN{Masahiro Sakai}
\IEEEauthorblockA{
Email: masahiro.sakai@gmail.com}
}

\maketitle

\begin{abstract}
PRINTEMPS is a metaheuristic solver originally developed for general integer linear programming (ILP) problems. For the MaxSAT Evaluation 2026, it has been extended to handle MaxSAT problems by linearizing each WCNF clause into a 0--1 inequality and translating soft clauses into a weighted objective penalty.
\end{abstract}

\section{Introduction}
PRINTEMPS (PoRtable INTEger Mathematical Programming Solver) is a metaheuristic solver developed by the first author for general integer linear programming (ILP) problems \cite{printemps}.
For the Pseudo-Boolean Competition 2025 \cite{pb25}, it was extended to handle Pseudo-Boolean (PB) problems, including those with nonlinear terms and soft constraints \cite{printemps-pb25}.
The MaxSAT extension submitted here reuses the PB encoding machinery for soft constraints and adds a WCNF front-end together with an anytime-style output protocol.
This document describes the supported tracks, the computational overview of PRINTEMPS, and the MaxSAT-specific extension.

\section{Supported Tracks}
Since PRINTEMPS is based on heuristics, it can prove neither the optimality of a solution nor the unsatisfiability of an instance in general.
Therefore, we submitted to the {\bf Main Tracks for Anytime Solvers}, namely {\bf ANYTIME-W} (weighted) and {\bf ANYTIME-UW} (unweighted).
The same binary handles both tracks, since the only difference between them is the weight distribution of the soft clauses in the input WCNF file \cite{mse26}.

The solver consumes the new WCNF format (post-2022): hard clauses are prefixed with \texttt{h}, soft clauses begin with an integer weight, and each clause is terminated with \texttt{0}.
Upon termination---either at the wall-clock limit or after receiving \texttt{SIGTERM} from the judging harness---the solver outputs the best assignment found so far as an \texttt{o}/\texttt{v} pair followed by a single status line \texttt{s SATISFIABLE} (exit code 10) or \texttt{s UNKNOWN} (exit code 0).
As an additional safety net, the wrapper also reports \texttt{s UNSATISFIABLE} (exit code 20) when bound propagation during preprocess detects that the hard clauses are infeasible.

\section{Computational Overview of PRINTEMPS}
An ILP problem can be formulated as follows:
%
\begin{IEEEeqnarray}{rcl}
    (\mathrm{ILP}): &\underset{\bm{x} \in \mathbb{Z}^{N},\, \bm{l} \le \bm{x} \le \bm{u}}{\mathrm{minimize}}\,& \bm{c}^{\top}\bm{x} \nonumber \\
    &\mathrm{subject\,to} \enspace & \bm{A}_{1}^{\top}\bm{x} = \bm{b}_{1}, \enspace \bm{A}_{2}^{\top}\bm{x} \le \bm{b}_{2}, \nonumber
\end{IEEEeqnarray}
%
where $N$ is the number of variables, $\bm{c} \in \mathbb{R}^{N}$ is the cost vector, and $\bm{l}, \bm{u} \in \mathbb{Z}^{N}$ $(\bm{l} \le \bm{u})$ are the lower and upper bounds of $\bm{x}$, respectively.
The matrices $\bm{A}_{1}, \bm{A}_{2}$ and the vectors $\bm{b}_{1}, \bm{b}_{2}$ have appropriate dimensions consistent with $\bm{x}$.

PRINTEMPS searches for solutions to an instance of (ILP) using Weighted Tabu Search that minimizes an objective function penalized by constraint violations \cite{Nonobe.1998}.
It also incorporates performance-enhancing techniques tailored for (ILP).
They include instance size reduction, dependent variable extraction, neighborhood filtering, and incremental penalty evaluation \cite{Koguma.2024}.

\section{Extension for MaxSAT Problems}
\subsection{Encoding of clauses}
Let $x_{i} \in \{0, 1\}$ denote the propositional variable indexed by $i$, and let a clause be a disjunction of literals $C = (\ell_{1} \vee \ell_{2} \vee \cdots \vee \ell_{k})$.
We split the literals into positive ones $P(C) = \{i \mid x_{i} \in C\}$ and negative ones $N(C) = \{i \mid \neg x_{i} \in C\}$.
The clause is satisfied if and only if at least one positive literal evaluates to $1$ or at least one negated variable evaluates to $0$, which is equivalent to
%
\begin{IEEEeqnarray}{c}
    \sum_{i \in P(C)} x_{i} - \sum_{j \in N(C)} x_{j} \;\ge\; 1 - |N(C)|. \nonumber
\end{IEEEeqnarray}
%
Each hard clause is added to the model directly as such a linear constraint.

\subsection{Soft clauses}
A soft clause with weight $w \ge 1$ is handled as a special case of the soft inequality treatment introduced for PB problems \cite{printemps-pb25}.
Since a clause is a one-sided $\ge$ constraint, only a single lower-side slack variable $y^{-} \in \{0, 1\}$ is required:
%
\begin{IEEEeqnarray}{c}
    \sum_{i \in P(C)} x_{i} - \sum_{j \in N(C)} x_{j} + y^{-} \;\ge\; 1 - |N(C)|, \nonumber
\end{IEEEeqnarray}
%
together with the penalty $w \cdot y^{-}$ added to the objective function.
Because the unrelaxed left-hand side falls below the right-hand side by exactly one when the clause is falsified, $y^{-}$ acts as the indicator of clause falsification, and the total objective penalty equals the sum of weights of unsatisfied soft clauses.

Soft-clause weights are read from the WCNF file as 64-bit unsigned integers and converted to double-precision floating-point coefficients in the objective penalty.
Weight sums up to $2^{53}$ are represented exactly; beyond that, rounding may occur but does not affect feasibility of the underlying assignment.

\subsection{Anytime behavior}
The MaxSAT Evaluation anytime tracks score solvers based on the quality of every reported intermediate solution \cite{mse26}.
The wrapper subscribes to the per-iteration callback of PRINTEMPS' Tabu Search loop and emits a fresh \texttt{o}/\texttt{v} pair whenever the current feasible incumbent strictly improves over the previously reported one.
The tracks also mandate single-threaded execution; PRINTEMPS is therefore configured to use a single OpenMP thread.

\section{Provenance}
The submitted system extends PRINTEMPS \cite{printemps}, an open-source C++ metaheuristic ILP solver originally authored by the first author. PRINTEMPS is distributed as a header-only library and has no proprietary runtime dependencies.
The PRINTEMPS core (the ILP modeler/solver library and its Weighted Tabu Search algorithm) and the Pseudo-Boolean Competition 2025 extension were both developed without generative AI assistance.
The structural layout of the competition wrapper and the soft-constraint encoding pattern are inherited from our Pseudo-Boolean Competition 2025 submission \cite{printemps-pb25}: the new wrapper under \texttt{extra/maxsat\_evaluation/} directly mirrors the previous \texttt{extra/pb\_competition\_2025/}.

The components newly added for this submission, with approximate line counts, are:
\begin{itemize}
\item the WCNF parser module \texttt{printemps/wcnf/} ($\sim$330 lines);
\item the \texttt{import\_wcnf} method on the IP model in \texttt{printemps/model/model.h} ($\sim$135 lines);
\item the competition wrapper under \texttt{extra/maxsat\_evaluation/} ($\sim$340 lines);
\item unit tests for the parser and an end-to-end test for the wrapper ($\sim$420 lines).
\end{itemize}
This new code, together with the corresponding build-system updates and the present description, was developed with the assistance of Claude Code \cite{claude-code} (Anthropic, invoked with the Claude Opus 4.7 model) used in a conversational pair-programming style. Every generated change was reviewed by the second author before being committed; the second author also directed the design, including the decision to keep the wrapper isolated under \texttt{extra/}, the WCNF-to-IP encoding scheme, and the error-handling protocol for the anytime tracks. The competition's regression test suite \cite{mse26} was used during development to validate parser correctness and the \texttt{s UNSATISFIABLE} exit path triggered by bound-propagation infeasibility.

The submitted Linux binary is built statically inside the project's Docker image (Ubuntu~22.04) and contains no proprietary or closed-source components.

\section{Conclusion}
The metaheuristic ILP solver PRINTEMPS has been extended to handle MaxSAT problems by reusing the PB soft-constraint encoding and adding a WCNF front-end and an anytime output protocol.

\begin{thebibliography}{1}
\bibitem{printemps}
PRINTEMPS C++ metaheuristics modeler/solver for general integer optimization problems, \url{https://snowberryfield.github.io/printemps/}.

\bibitem{pb25}
Pseudo-Boolean Competition 2025, \url{https://www.cril.univ-artois.fr/PB25/}.

\bibitem{printemps-pb25}
Y.~Koguma and M.~Sakai, ``An Incomplete Approach to Pseudo-Boolean Problems with Metaheuristic Solver PRINTEMPS,'' Solver description, Pseudo-Boolean Competition 2025, \url{https://www.cril.univ-artois.fr/PB25/descr/PRINTEMPS-Pseudo_Boolean_Competition_2025.pdf}.

\bibitem{mse26}
MaxSAT Evaluation 2026, \url{https://maxsat-evaluations.github.io/2026/}.

\bibitem{Nonobe.1998}
K.~Nonobe and T.~Ibaraki, ``A Tabu Search Approach to the Constraint Satisfaction Problem as a General Problem Solver,'' \emph{Eur. J. Oper. Res.}, Vol.~106, pp.~599--623, 1998, doi:10.1016/S0377-2217(97)00294-4.

\bibitem{Koguma.2024}
Y.~Koguma, ``Tabu Search-Based Heuristic Solver for General Integer Linear Programming Problems,'' {\em IEEE Access}, Vol.~12, pp.~19059--19076, 2024, doi: 10.1109/ACCESS.2024.3361323.

\bibitem{claude-code}
Anthropic, ``Claude Code,'' an AI-assisted coding agent, \url{https://www.anthropic.com/claude-code}.

\end{thebibliography}
\end{document}
